---
title: 'Homework: Numerical linear algebra'
output:
  html_document:
    df_print: paged
  html_notebook:
    code_folding: none
    highlight: textmate
---

<script type="text/x-mathjax-config">
MathJax.Hub.Config({
  TeX: {
    Macros: {
      bA: "{\\boldsymbol{A}}",
      bx: "{\\boldsymbol{x}}",
      bb: "{\\boldsymbol{b}}"
    }
  }
});
</script>


```{r setup, include=FALSE}
source(file.path("..", "R", "util.R"))

required_packages <- c("microbenchmark")
install_and_load_packages(required_packages)
```


# Exercise 1: Comparing different numerical linear algebra algorithms for solving linear systems

In this exercise, we consider the problem of solving a linear system $\bA \bx = \bb$ for $\bx$.
We compare the three methods we learned in the class: LU, Cholesky, and QR decompositions.
(Of course, LU applies to more general systems and QR can be used to solve least squares, but here we focus on positive definite systems.)

## Part A: Racing for solution &mdash; speed comparison 

We first compare their computational speed. 
Fill in the code below and `bench::mark()` the three algorithms.

**Questions:**
What are relative speeds among the algorithms?
Do relative speeds agree with what you expect from the complexity analyses?
If not (quite), why might that be?

Based on the comparison results using median execution time, the fastest algorithm is solve (using LU decompostion), followed by solve via Cholesky and QR decomposition. 

The relative speeds are not under expectation. Since the overall cost of solving $Ax=b$ via Cholesky is $O(\frac{1}{3}p^3)$, via QR decomp is $O(2np^2)$ and solve a linear system (LU) may requires $O(\frac{2}{3}p^2)$. However, solve function is faster than solving $Ax=b$ via Cholesky. The speed from fastest to slowest is solve (LU) > Cholesky > QR decomposition. Since QR decomp is $O(2np^2)$ cost, it is fair to be the slowest. However, the solve (LU) is faster than Cholesky may be due to that during solve function (LU), only A and b are input into LAPACK routine function, however, we store Cholesky output and R function is not as fast as directly using LAPACK routine library. 

**Note:**
I misspoke about the R's built-in `chol()` function during the lecture:
when applied to a positive-definite `A`, the function actually returns an _upper-triangular_ matrix `R` such that `t(R) %*% R == A`.

```{r, eval=T}
# Import the `rand_positive_def_matrix` function
source(file.path("R", "random_matrix.R"))

mat_size <- 1024L
A <- rand_positive_def_matrix(mat_size, cond_num = 1e3)
b <- rnorm(mat_size)

solve_via_cholesky <- function(A, b) {
  R = chol(A)
  x = backsolve(R, forwardsolve(R, b, upper.tri = T, transpose = T))
  return(x)
}

bench::mark(
  solve(A, b)
)
bench::mark(
  solve_via_cholesky(A, b)
)
bench::mark(
  # Solve via QR decomposition
  qr.solve(A, b)
)
```

## Part B: Competition in terms of numerical accuracy/stability

We now compare the three methods in terms of numerical accuracy/stability.
To this end, we set up the following simulation study. 
We first generate a "ground truth" solution vector $\bx_0$.
We then compute an "numerical-error-free" $\bb = \bA \bx_0$ by carrying out the matrix-vector multiplication using the `long double` type, which (on most hardware and compilers) provides [additional 12 bits of precision](https://en.wikipedia.org/wiki/Extended_precision#x86_extended_precision_format).
Of course, the vector $\bb$ computed as such still suffers from numerical errors, but the idea is that the numerical errors from this high-precision matrix-vector multiplication is much smaller than the errors caused by numerically solving $\bA \bx = \bb$ for $\bx$.
We can thus assess the accuracy of the three solvers by comparing the numerically computed $\bx$ to the ground truth $\bx_0$.

### Task &#x2F00;

First compare the outputs of matrix-vector multiplication $\bx \to \bA \bx$ using `double` and `long double` using the provided Rcpp functions.

**Questions:**
What is the relative difference in $\ell^2$-norm? 
How about the coordinate-wise relative differences?
Are the observed magnitudes of the differences what you'd expect?

The relative difference in $\ell^2$-norm is 1.185409e-15, the summary of the coordinate-wise relative differences is as follows. Since we use long double, additional 12 bits of precision, using two ways to calculate the same thing will give a different result, and based on the machine precision, it reflects at the level of around $10^{-15}$ and $10^{-16}$.

```{r, eval=T}
Rcpp::sourceCpp(file.path("src", "matvec_double.cpp"))
Rcpp::sourceCpp(file.path("src", "matvec_ldouble.cpp"))

set.seed(1918)

mat_size <- 1024L
A <- rand_positive_def_matrix(mat_size, cond_num = 1e3)
x <- rnorm(mat_size)

# Fill in
double_result = matvec_double(A,x)
ldouble_result = matvec_ldouble(A,x)

sqrt(sum((double_result - ldouble_result)^2)/sum(ldouble_result^2))

summary((double_result - ldouble_result)/ldouble_result)
```

### Task &#x2F06;

Now randomly generate $\bA$ so that its condition number is $10^6$.
Then solve a positive-definite system $\bA \bx = \bb$ for $\bx$ using the three algorithms and compare their outputs to the ground truth $\bx_0$.

**Questions:**
Which algorithm appears to be more accurate than the others? 
Visually demonstrate your answer.

From the plot, we could see the solve is more accurate than cholesky decompostion, the worst is QR decomposition.

```{r, eval=T}
set.seed(1918)
cond_num <- 1e6

# Larger matrices could incur substantial computational time under base R BLAS
mat_size <- 1024L 

A <- rand_positive_def_matrix(mat_size, cond_num)
x <- rnorm(mat_size)

# Fill in
b = matvec_ldouble(A,x)

solve_diff = solve(A, b) - x

cholesky_diff = solve_via_cholesky(A, b)- x

qr_diff = qr.solve(A, b) - x

library(ggplot2)
data_plot = data.frame(diff=c(solve_diff,cholesky_diff,qr_diff),
                       x=x,
                       algorithm = c(rep("solve",mat_size),
                                     rep("cholesky",mat_size),
                                     rep("QR",mat_size)))
ggplot(data = data_plot, aes(x=x,y=diff,group = algorithm,color=algorithm))+geom_point()

```

### Task &#x4E09;

In Task &#x2F06;, we compared the three algorithms in one randomly generated example.
Now we consider a more systematic (though hardly comprehensive) comparison via repeated simulations.
We also vary the condition number of $\bA$ and assess whether the results would hold across varying degrees of ill-conditioning.

**Questions/To-do's:**

* Using the starter code provided, calculate various summary measures of the numerical errors.
* Integrate into the provided code one another (or more, if you like) meaningful metric(s) of your choice to summarize the numerical error.
* Visually explore how the three algorithms compare with each other in their accuracy. See if different error metrics tell different stories; they might or might not.
* Vary the condition number in the range $10^6 \sim 10^{12}$, e.g. by trying $10^6$, $10^9$, and $10^{12}$.
* Do you see any patterns in the numerical errors across the three algorithms, metrics, and/or condition numbers? Show some plots to support your conclusion.

(1) Within the same condition numbers, different metrics give similar results when comparing three algorithms, and we can see QR has a largest relative error, solve function and choleksy decomposition give a similar result, but choleksy decomposition may give a slightly larger error than solve function.

(2) For larger condition number, all three algorithms have larger relative error comparing with the smaller condition number, this can be seen based on the magnitude of the y-axis. 


**Note:** 
The QR solver will throw an error when the system is ill-condition enough that the numerical solution might not be very accurate. 
To force it to return the solution in any case, set `tol = .Machine$double.eps`.

```{r, eval=T}
# Utility functions for bookkeeping simulation results.
source(file.path("R", "num_linalg_sim_study_helper.R"))

n_sim <- 32L
mat_size <- 512L
cond_num <- 1e6
metrics <- c("norm", "median", "five_percentile", "ninety_five_percentile", "average")
  # TODO: add another metric and modify the helper script accordingly

rel_error_list <- lapply(
  c("lu", "chol", "qr"), 
  function(method) pre_allocate_error_list(n_sim, metrics)
)

for (sim_index in 1:n_sim) {
  A <- rand_positive_def_matrix(mat_size, cond_num)
  x <- rnorm(mat_size) 
  b <- matvec_ldouble(A, x)
  x_approx <- list( 
    lu = solve(A, b),
    chol = solve_via_cholesky(A, b),
    qr = qr.solve(A, b)
  )
  for (method in c("lu", "chol", "qr")) {
    rel_error <- lapply(
      metrics, 
      function (metric) calc_rel_error(x, x_approx[[method]], metric)
    )
    names(rel_error) <- metrics
    for (metric in names(rel_error)) {
      rel_error_list[[method]][[metric]][sim_index] <- rel_error[[metric]]
    }
  }
}

# TODO: visually compare errors
plot_compare <- function(metric){
  data_plot = data.frame(value=c(rel_error_list[["lu"]][[metric]],
                               rel_error_list[["chol"]][[metric]],
                               rel_error_list[["qr"]][[metric]]),
                       id=rep(seq(1,n_sim,1),3),
                       algorithm = c(rep("lu",n_sim),
                                     rep("chol",n_sim),
                                     rep("qr",n_sim)))
 p = ggplot(data = data_plot, 
            aes(x=id,y=value,group =algorithm,color=algorithm))+geom_point()
 return(p)
}

plot_compare("norm")
plot_compare("median")
plot_compare("five_percentile")
plot_compare("ninety_five_percentile")
plot_compare("average")

```

```{r, eval=T}
# Utility functions for bookkeeping simulation results.
source(file.path("R", "num_linalg_sim_study_helper.R"))

n_sim <- 32L
mat_size <- 512L
cond_num <- 1e9
metrics <- c("norm", "median", "five_percentile", "ninety_five_percentile", "average")
  # TODO: add another metric and modify the helper script accordingly

rel_error_list <- lapply(
  c("lu", "chol", "qr"), 
  function(method) pre_allocate_error_list(n_sim, metrics)
)

for (sim_index in 1:n_sim) {
  A <- rand_positive_def_matrix(mat_size, cond_num)
  x <- rnorm(mat_size) 
  b <- matvec_ldouble(A, x)
  x_approx <- list( 
    lu = solve(A, b),
    chol = solve_via_cholesky(A, b),
    qr = qr.solve(A, b, tol = .Machine$double.eps)
  )
  for (method in c("lu", "chol", "qr")) {
    rel_error <- lapply(
      metrics, 
      function (metric) calc_rel_error(x, x_approx[[method]], metric)
    )
    names(rel_error) <- metrics
    for (metric in names(rel_error)) {
      rel_error_list[[method]][[metric]][sim_index] <- rel_error[[metric]]
    }
  }
}

# TODO: visually compare errors
plot_compare <- function(metric){
  data_plot = data.frame(value=c(rel_error_list[["lu"]][[metric]],
                               rel_error_list[["chol"]][[metric]],
                               rel_error_list[["qr"]][[metric]]),
                       id=rep(seq(1,n_sim,1),3),
                       algorithm = c(rep("lu",n_sim),
                                     rep("chol",n_sim),
                                     rep("qr",n_sim)))
 p = ggplot(data = data_plot, 
            aes(x=id,y=value,group =algorithm,color=algorithm))+geom_point()
 return(p)
}

plot_compare("norm")
plot_compare("median")
plot_compare("five_percentile")
plot_compare("ninety_five_percentile")
plot_compare("average")

```

```{r, eval=T}
# Utility functions for bookkeeping simulation results.
source(file.path("R", "num_linalg_sim_study_helper.R"))

n_sim <- 32L
mat_size <- 512L
cond_num <- 1e12
metrics <- c("norm", "median", "five_percentile", "ninety_five_percentile", "average")
  # TODO: add another metric and modify the helper script accordingly

rel_error_list <- lapply(
  c("lu", "chol", "qr"), 
  function(method) pre_allocate_error_list(n_sim, metrics)
)

for (sim_index in 1:n_sim) {
  A <- rand_positive_def_matrix(mat_size, cond_num)
  x <- rnorm(mat_size) 
  b <- matvec_ldouble(A, x)
  x_approx <- list( 
    lu = solve(A, b),
    chol = solve_via_cholesky(A, b),
    qr = qr.solve(A, b, tol = .Machine$double.eps)
  )
  for (method in c("lu", "chol", "qr")) {
    rel_error <- lapply(
      metrics, 
      function (metric) calc_rel_error(x, x_approx[[method]], metric)
    )
    names(rel_error) <- metrics
    for (metric in names(rel_error)) {
      rel_error_list[[method]][[metric]][sim_index] <- rel_error[[metric]]
    }
  }
}

# TODO: visually compare errors
plot_compare <- function(metric){
  data_plot = data.frame(value=c(rel_error_list[["lu"]][[metric]],
                               rel_error_list[["chol"]][[metric]],
                               rel_error_list[["qr"]][[metric]]),
                       id=rep(seq(1,n_sim,1),3),
                       algorithm = c(rep("lu",n_sim),
                                     rep("chol",n_sim),
                                     rep("qr",n_sim)))
 p = ggplot(data = data_plot, 
            aes(x=id,y=value,group =algorithm,color=algorithm))+geom_point()
 return(p)
}

plot_compare("norm")
plot_compare("median")
plot_compare("five_percentile")
plot_compare("ninety_five_percentile")
plot_compare("average")

```