\documentclass{article}


\title{One reason to version control source files only}

\begin{document}

\maketitle

This is the first sentence.

This is the sentence number 2.
This is the sentence number 3.
This is the sentence number 4.
This is the sentence number 5.
This is the sentence number 6.
This is the sentence number 7.
This is the sentence number 8.
This is the sentence number 9.
This is the sentence number 10.
This is the sentence number 11.
This is the sentence number 12.
This is the sentence number 13.
This is the sentence number 14.
This is the sentence number 15.
This is the sentence number 16.
This is the sentence number 17.
This is the sentence number 18.
This is the sentence number 19.
This is the sentence number 20.
This is the sentence number 21.
This is the sentence number 22.
This is the sentence number 23.
This is the sentence number 24.
This is the sentence number 25.
This is the sentence number 26.
This is the sentence number 27.
This is the sentence number 28.
This is the sentence number 29.
This is the sentence number 30.
This is the sentence number 31.
This is the sentence number 32.
This is the sentence number 33.
This is the sentence number 34.
This is the sentence number 35.
This is the sentence number 36.
This is the sentence number 37.
This is the sentence number 38.
This is the sentence number 39.
This is the sentence number 40.
This is the sentence number 41.
This is the sentence number 42.
This is the sentence number 43.
This is the sentence number 44.
This is the sentence number 45.
This is the sentence number 46.
This is the sentence number 47.
This is the sentence number 48.
This is the sentence number 49.
This is the sentence number 50.
This is the sentence number 51.
This is the sentence number 52.
This is the sentence number 53.
This is the sentence number 54.
This is the sentence number 55.
This is the sentence number 56.
This is the sentence number 57.
This is the sentence number 58.
This is the sentence number 59.
This is the sentence number 60.
This is the sentence number 61.
This is the sentence number 62.
This is the sentence number 63.
This is the sentence number 64.
This is the sentence number 65.
This is the sentence number 66.
This is the sentence number 67.
This is the sentence number 68.
This is the sentence number 69.
This is the sentence number 70.
This is the sentence number 71.
This is the sentence number 72.
This is the sentence number 73.
This is the sentence number 74.
This is the sentence number 75.
This is the sentence number 76.
This is the sentence number 77.
This is the sentence number 78.
This is the sentence number 79.
This is the sentence number 80.
This is the sentence number 81.
This is the sentence number 82.
This is the sentence number 83.
This is the sentence number 84.
This is the sentence number 85.
This is the sentence number 86.
This is the sentence number 87.
This is the sentence number 88.
This is the sentence number 89.
This is the sentence number 90.
This is the sentence number 91.
This is the sentence number 92.
This is the sentence number 93.
This is the sentence number 94.
This is the sentence number 95.
This is the sentence number 96.
This is the sentence number 97.
This is the sentence number 98.
This is the sentence number 99.
This is the sentence number 100.
\end{document}